Sea, Lake, and Overland Surges from Hurricanes \citep{jelesnianski1992} is a computer model developed by the National Weather Service to simulate storm surge, and its associated inundation caused by hurricanes.   Given storm characteristics, the model takes into account local topology, bathymetry, and surge management devices such as levees, to generate a spatial field of inundation---the maximum observed height of water above ground level (or above normal water level for a data point in a body of water) for the duration of the storm at a location.   These storm characteristics include data pertaining to the eye of the storm when it made landfall---bearing, velocity, and latitude, minimum atmospheric pressure of the storm when it made landfall, and projections of sea level rise over time.  A \emph{simulation} from the model is a grid containing \num{23119800} elements wih a spatial resolution of \num{0.001} degrees \makenote{(verify)}, covering an area extending from Virginia Beach, Virginia, to Long Island, New York.  We have \num{4000} such simulations, produced by SLOSH from a sample of storm characteristics.

\makenote{Insert graphic of coverage of model}

Inundation can be catastrophic.  Setting aside the potential for loss of life, flooding can impose costly damage to property: inundation of homes and businesses destroys posessions, and damages buildings through saturating walls and eroding foundations., and thus the risk of damage from flooding provides incentive for flood insurance. \makenote{expand}\needcite  Flooding can damage vehicles, and a single storm can force insurance companies to declare large quantities of vehicles as total losses. \makenote{expand}\needcite.  Flooding damages agriculture: beyond destruction of currently growing or stored crops, or the drowning of livestock, inundation by storm surge results in the ground absorbing salt, destroying production capacity of the field until abatement. Flooding can damage infrastucture: flooded roads can be washed out or have their foundation damaged, flooded sewers and sewage treatment plants can release their contents above ground imposing additional environmental costs.  Flooded power infrastructure, such as transformers can short out causing additional damage. \citep{hutchings2021}.  Inundation can impose additional burdens in the moment:  inundation negatively affects the quality of emergency services, such as a hospital being rendered unable to intake patients.  Sufficient flooding may even render a provider entirely out of commission. 

It makes sense to consider

The projected gamma based model is not an inherently spatial model.  As such, rather than holding the entire grid in memory, it makes sense to consider data pertaining only to landmarks of interest.    If one is performing contingency planning in preparation for the storm season, it would be helpful to know the probability of such services being rendered inoperable simultaneously, precisely when they are needed most.  If one is seeking sites for a new service provider, it would be helpful understand the likelihood of a proposed location being rendered inoperable in the same manner.  For these reasons, and considering the computational complexity of a 23 million dimensional model, rather than consider the entire grid of data pertaining to any particular storm, we subset the data to grid cells in the vicinity of such locations of interest.  These locations are gathered from the 2023 US Census Bureau's \emph{TIGER} database \needcite and specifically the point and landmark file, and only grid cells that experienced \emph{some} inundation in greater than 2 \makenote{(subject to change)} percent of storm simulations.   This last requirement arises as consequence of fitting the generalized Pareto: the threshold for excesses $b$ needs to be set above the minimum observed value $(0)$ of the dataset.  \makenote{(idea: zero-inflated projected gamma?)}  We identify grid cells in the vicinity of such locations by establishing a buffer 







